---
title: "Data Science I: HW2"
author: "Sydney Musick"
output: github_document
---

# Problem 1

### Packages upload.
```{r setup, echo = FALSE}
library(readxl)
library(tidyverse)
library(lubridate)
```

### Data upload and cleaning for trash data.
```{r trashwheel upload, echo = FALSE}
trashwheel_df = 
  drop_na(
    janitor::clean_names(
          read_excel("trashwheels.xlsx", sheet = 1, col_names = TRUE, range = "A2:N534")
         )) %>%
  mutate(sports_balls = round(sports_balls, digits = 0))
trashwheel_df
```

### Data upload and cleaning for 2018 precipitation data.
```{r 2018 upload, echo = FALSE}
precipitation_2018_df = 
  drop_na(
    janitor::clean_names(
          read_excel("trashwheels.xlsx", sheet = 7, col_names = TRUE, range = "A2:B14")
         )) %>%
  mutate(year = 2018)
precipitation_2018_df
```

### Data upload and cleaning for 2019 precipitation data.
```{r 2019 upload, echo = FALSE}
precipitation_2019_df = 
  drop_na(
    janitor::clean_names(
          read_excel("trashwheels.xlsx", sheet = 6, col_names = TRUE, range = "A2:B14")
         )) %>%
  mutate(year = 2019)
precipitation_2018_df
```

### Combining the 2018 and 2019 dataframes.
```{r combine, echo = FALSE}
precipitation_df = 
  bind_rows(precipitation_2018_df, precipitation_2019_df) %>%
  mutate(month = c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December", "January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))
precipitation_df
```

The `trashwheel_df` dataset has `r nrow(trashwheel_df)` observations. It includes variables `r names(trashwheel_df)`. The median number of sports balls in a dumpster in 2019 was `r median(pull(trashwheel_df, sports_balls))`. 

The `precipitation_df` dataset has `r nrow(precipitation_df)` observations. It includes variables `r names(precipitation_df)`. The total precipitation in 2018 was `r sum(pull(precipitation_2018_df, total))` inches.


# Problem 2

### Data upload and cleaning for pols data.
```{r pols upload, echo = FALSE}
pols_df = 
  janitor::clean_names(
          read_csv("fivethirtyeight_datasets/pols-month.csv", col_names = TRUE)) %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = replace(month, month == "01", "January")) %>%
  mutate(month = replace(month, month == "02", "February")) %>%
  mutate(month = replace(month, month == "03", "March")) %>%
  mutate(month = replace(month, month == "04", "April")) %>%
  mutate(month = replace(month, month == "05", "May")) %>%
  mutate(month = replace(month, month == "06", "June")) %>%
  mutate(month = replace(month, month == "07", "July")) %>%
  mutate(month = replace(month, month == "08", "August")) %>%
  mutate(month = replace(month, month == "09", "September")) %>%
  mutate(month = replace(month, month == "10", "October")) %>%
  mutate(month = replace(month, month == "11", "November")) %>%
  mutate(month = replace(month, month == "12", "December")) %>%
  mutate(prez_gop = replace(prez_gop, prez_gop == "1", 11)) %>%
  mutate(prez_dem = replace(prez_dem, prez_dem == "1", 99)) %>%
  mutate(president = prez_gop + prez_dem ) %>%
  mutate(president = replace(president, president == 11, "gop")) %>%
  mutate(president = replace(president, president == 99, "dem")) %>%
  select(-day, -prez_gop, -prez_dem) %>%
  mutate(year = as.numeric(year))
pols_df
```

### Data upload and cleaning for snp data.
```{r snp upload, echo = FALSE}
snp_df = 
  janitor::clean_names(
  read_csv("fivethirtyeight_datasets/snp.csv", col_names = TRUE)) %>%
  separate(date, into = c("month", "day", "year")) %>%
  mutate(month = replace(month, month == "1", "January")) %>%
  mutate(month = replace(month, month == "2", "February")) %>%
  mutate(month = replace(month, month == "3", "March")) %>%
  mutate(month = replace(month, month == "4", "April")) %>%
  mutate(month = replace(month, month == "5", "May")) %>%
  mutate(month = replace(month, month == "6", "June")) %>%
  mutate(month = replace(month, month == "7", "July")) %>%
  mutate(month = replace(month, month == "8", "August")) %>%
  mutate(month = replace(month, month == "9", "September")) %>%
  mutate(month = replace(month, month == "10", "October")) %>%
  mutate(month = replace(month, month == "11", "November")) %>%
  mutate(month = replace(month, month == "12", "December")) %>%
  select (-day) %>%
  select(year, month, close) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(year = ifelse(68 > year %% 100 & 16 > year %% 100, 2000 + year %% 100, 1900 + year %% 100))
snp_df
```

### Data upload and cleaning for unemployment data.
```{r unemployment upload, echo = FALSE}
unemployment_df = 
  janitor::clean_names(
  read_csv("fivethirtyeight_datasets/unemployment.csv", col_names = TRUE)) %>%
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "unemployment_rate") %>%
  mutate(month = replace(month, month == "jan", "January")) %>%
  mutate(month = replace(month, month == "feb", "February")) %>%
  mutate(month = replace(month, month == "mar", "March")) %>%
  mutate(month = replace(month, month == "apr", "April")) %>%
  mutate(month = replace(month, month == "may", "May")) %>%
  mutate(month = replace(month, month == "jun", "June")) %>%
  mutate(month = replace(month, month == "jul", "July")) %>%
  mutate(month = replace(month, month == "aug", "August")) %>%
  mutate(month = replace(month, month == "sep", "September")) %>%
  mutate(month = replace(month, month == "oct", "October")) %>%
  mutate(month = replace(month, month == "nov", "November")) %>%
  mutate(month = replace(month, month == "dec", "December"))
unemployment_df
```

### Combining the pols and snp dataframes.
```{r combine snp pols, echo = FALSE}
pols_snp_df =
  inner_join(pols_df, snp_df, by = c("year", "month"))
pols_snp_df
```

### Combining the pols_snp and unemployment dataframes.
```{r combine snp_pols unemployment, echo = FALSE}
all_df =
  inner_join(pols_snp_df, unemployment_df, by = c("year", "month"))
all_df
```

The `pols_df` dataset has `r nrow(pols_df)` rows and `r ncol(pols_df)` columns. It includes variables `r names(pols_df)`. It contains data on the political alignment of the president, governors, senators, and representatives from 1947-2015.

The `snp_df` dataset has `r nrow(snp_df)` rows and `r ncol(snp_df)` columns. It includes variables `r names(snp_df)`. It contains data on the closing values of the S&P stock index from 1950-2015.

The `unemployment_df` dataset has `r nrow(unemployment_df)` rows and `r ncol(unemployment_df)` columns. It includes variables `r names(unemployment_df)`. It contains data on the unemployment rate from 1948-2015.

The `all_df` dataset is a merging of dataframes `pols_df`, `snp_df`, and `unemployment_df`. It has `r nrow(all_df)` rows and `r ncol(all_df)` columns. It includes variables `r names(all_df)`. It contains data on the political alignment of the president, governors, senators, and representatives, the closing values of the S&P stock index, and the unemployment rate from 1950-2015.

# Problem 3

### Uploading baby name file.
```{r baby name upload, echo = FALSE}
names_df = 
  janitor::clean_names(
  read_csv("Popular_Baby_Names.csv", col_names = TRUE)) %>%
  mutate(ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER")) %>%
  mutate(ethnicity = replace(ethnicity, ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC")) %>%
  mutate(ethnicity = replace(ethnicity, ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC")) %>%
  mutate(childs_first_name = toupper(childs_first_name)) %>%
  distinct()
```

### Making a table for the popularity of name "Olivia".
```{r olivia table, echo = FALSE}
names_df %>%
  filter(childs_first_name == "OLIVIA") %>%
  select(ethnicity, year_of_birth, rank) %>%
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = "rank")
```

### Making a table for the popularity of name "Jayden".
```{r jayden table, echo = FALSE}
names_df %>%
  filter(childs_first_name == "JAYDEN") %>%
  select(ethnicity, year_of_birth, rank) %>%
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = "rank")
```

### Plotting count against rank for male, white non-hispanic children born in 2016.
```{r plot, echo = FALSE}
names_df %>%
  filter(ethnicity == "WHITE NON HISPANIC") %>%
  filter(gender == "MALE") %>%
  filter(year_of_birth == "2016") %>%
  ggplot(aes(x = rank, y = count)) + geom_point() 
  ggsave("scatter_plot.pdf", height = 4, width = 6)
```


